---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(jstools)
```

# jstools

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![Travis build status](https://travis-ci.org/dreamRs/jstools.svg?branch=master)](https://travis-ci.org/dreamRs/jstools)
[![R build status](https://github.com/dreamRs/jstools/workflows/R-CMD-check/badge.svg)](https://github.com/dreamRs/jstools/actions)
<!-- badges: end -->

> Set of tools to work with JavaScript and CSS (and addins to interactively use them in RStudio)

Disclaimer: I use this in my personnal workflow when working with JavaScript or CSS, some functionnalities are similar to [{js}](https://github.com/jeroen/js) or [{reactR}](https://github.com/react-R/reactR).


## Installation

You can install the development version from GitHub with:

``` r
# install.packages("remotes")
remotes::install_github("dreamRs/jstools")
```


## Prettier : format code

Via https://prettier.io/

`prettier_js()` allow to reformat JavaScript code:

```{r, eval=FALSE}
prettier_js("function add(first, second) { return first + second; }")
```
```{r, echo=FALSE, comment=""}
cat(prettier_js("function add(first, second) { return first + second; }"))
```

`pettier_css()` does the same thing for CSS:

```{r, eval=FALSE}
prettier_css("b{font-weight: bold;color:red;}")
```
```{r, echo=FALSE, comment=""}
cat(prettier_css("b{font-weight: bold;color:red;}"))
```



You can reformat a file with:

```{r, eval=FALSE}
prettier_file(input = "path/to/file.js", output = "path/to/reformated.js")
```

In Rstudio, you can use addin "Prettier" to reformat the current script (careful, it will overwrite it).



## Terser : JavaScript compressor

Via https://terser.org/

`terser()` can be used to compress JavaScript code :

```{r}
terser("function add(first, second) { return first + second; }")
```

You can compress a file (or several) with:

```{r, eval=FALSE}
terser_file(input = "path/to/file.js", output = "path/to/file.min.js")
terser_file(input = c("path/to/file1.js", "path/to/file2.js"), output = "path/to/file.min.js")
```

Source maps are crucial to the debugging process, thereby making it possible to reconstruct the original
JS code starting from a minified script. To correctly generate a source map, pass the __sourceMap__ option to `terser_options`. 
`includeSources = TRUE` is mandatory so that files may be retrieved by the web browser developer tools.
The folder will have the provided __root__ as name so it is important to name it consistently:

```{r}
path <- system.file("testfiles/htmlwidgets.js", package = "jstools")
# source maps options
o <- terser_options(
  sourceMap = list(
    root = "../../htmlwidgets-sources",
    filename = "htmlwidgets.min.js",
    url = "htmlwidgets.min.js.map",
    includeSources = TRUE
  )
)
# returns a list containing the minified code an the source map
terser_file(input = path, options = o)
```


## JSHint : code validation

Via https://jshint.com/

`jshint()` detect errors in JavaScript code:

```{r}
res <- jshint(c("var a = 2;", "var foo = 1;", "b = 3", "a + c"))
res$errors[, c("line", "reason")]
```

Use with a file : 

```{r}
path <- system.file("testfiles/example.js", package = "jstools")
jshint_file(input = path)
```

To validate a Shiny script, you can add `jQuery` and `Shiny` as global variables (same for `Htmlwidgets`:

```{r, eval=FALSE}
jshint_file(input = path, options = jshint_options(jquery = TRUE, globals = list("Shiny")))
```

You can check several scripts at once (for example custom Shiny input bindings in {shinyWidgets}):

```{r, eval=FALSE}
bindings <- list.files(
  path = system.file("assets", package = "shinyWidgets"),
  pattern = "bindings\\.js$",
  recursive = TRUE,
  full.names = TRUE
)
jshint_file(input = bindings, options = jshint_options(jquery = TRUE, globals = list("Shiny")))
```


In Rstudio, you can use addin "JSHint" to run code validation on current script.



## Babel : JavaScript compiler

Via https://babeljs.io/

`babel()` can be used to convert ES6+ code to ES5 to make it works in most browser:

```{r, eval=FALSE}
# Template literals are not supported in ES5
babel("var filename = `${Date.now()}.png`;")
```
```{r, echo=FALSE, comment=""}
cat(babel("var filename = `${Date.now()}.png`;"))
```



```{r, eval=FALSE}
# Arrow function neither
babel("[1, 2, 3].map((n) => n + 1);")
```
```{r, echo=FALSE, comment=""}
cat(babel("[1, 2, 3].map((n) => n + 1);"))
```



## Crass : CSS minifier and optimizer

Via https://github.com/mattbasta/crass

Minimize and optimize CSS code : 

```{r}
crass("b { font-weight: bold; }")
```


You can concatenate several files together:

```{r, eval=FALSE}
crass_file(
  input = c("path/to/file1.css", "path/to/file2.css"),
  output = "path/to/file.min.css"
)
```



## CSSO : CSS minifier with structural optimizations

Via https://github.com/css/csso

Merge identical rules together:

```{r}
csso(".foo { color: #ff0000; }\n.bar { color: rgba(255, 0, 0, 1); }")
```

You can also concatenate several files together:

```{r, eval=FALSE}
csso_file(
  input = c("path/to/file1.css", "path/to/file2.css"),
  output = "path/to/file.min.css"
)
```



